### 2004-07-12 Documentation/part2.tex
extended the \GLB command to handle a leading '@' (this is meta character for makeindex and thus has to be escaped by '"'). Added a label for section 7.2

If you want to have several experiments in one file, separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules and energy resolution
functions. All variables have their scope limited 
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}. The only exception are 
the counters defined by {\tt cross} and {\tt flux}. They are valid for 
the complete session, since one does not want too overload the flux and 
cross section information when using more than one experiment. This allows 
a consistent treatment of various experiments in one file.

As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include ./file\_1}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include NuFact.gls\\
        \#NEXT\#\\
        include JHFHK.gls
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. 
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments in the 
application software. In this case, {\tt \#CC} in {\tt NuFact.gls} has 
the value $0$ and in {\tt JHFHK.gls} it has the value $1$, 
which means that no shadowing occurs. Note that, alternatively, one can 
also load both files successively.

@@

In addition to the matter density profile, \GLOBES\ implements a simple treatment of matter density uncertainties by a scaling factor $\hat{\rho}$
of the matter density profile:
\begin{equation}
\label{eq:density_error}
\rho(x)=\hat{\rho} \cdot\rho_0(x)\,\left[\frac{\mathrm{g}}{\mathrm{cm}^3}\right] \, .
\end{equation}
Here $\rho_0(x)$ is the matter profile chosen by the user, and $\rho(x)$ is
the one used in the calculation of the oscillation probabilities. As it is described in \Sec~\ref{sec:externalinput}, the matter density uncertainty is implemented by using $\hat{\rho}$ as an additional oscillation parameter for each experiment, and by assuming to know $\hat{\rho}$ with an external precision. Thus, the central value of $\hat{\rho}$ (default value $1.0$) and its error $\delta \hat{\rho}$ have to be defined. The following code assumes a matter density uncertainty of $5\%$:\index{Density error}\index{Density center}
\begin{quote}
 {\tt \$densitycenter=1.0} \\
 {\tt \$densityerror=0.05}
\end{quote}
In most cases, this is in combination with a constant profile a very good approximation to take into account the realistic structure of the profile as well as its uncertainty~\cite{Ohlsson:2003ip}. 
 If one does not want to take into account  matter density uncertainties, 
one can either a very small value of $\hat{\rho}$, or the minimizer {\tt ChiNP} with $\hat{\rho}$ fixed in the application software.

@@

Before we come to the calculation algorithms, it is useful to understand
the general evaluation algorithm. As it is illustrated in \figu{energies}, 
\GLOBES\ uses several levels with respect to the energy ranges:
\begin{description}
\item[Sampling point level]
 This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points. The energy scale is the actual incident neutrino energy $E$. Note that the sampling range should exceed the energy bin range by about the energy resolution error, since the corresponding sampling points have to be evaluated by the algorithm. 
However, this exceeding is only
necessary to avoid aliasing effects, if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
For a manual definition of the sampling points (depends on algorithm), use
\begin{quote}
{\tt
\$sampling\_points = 20\\
\$sampling\_min =          4.0\\
\$sampling\_max =         50.0
}
\end{quote}
for equidistant sampling points. Arbitrarily spaced sampling points can 
be specified analogously to the energy bins below. 
Note that the algorithm expects to 
have at least as many sampling points as energy bins, since otherwise 
the number of (unphysical) sampling constrains the experiment performance.
\item[Bin level] This level is determined by the experiment. Note than energy
bin sizes much smaller than the energy resolution will not improve the results. The energy bin range and the number of energy bins do always have to be specified. For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about the energy calibration error in order to avoid aliasing effects.

@@

As far as the parameterization for the energy resolution function 
$R^c(E,E')$ in \eq~(\ref{eq:kernel}) is concerned, the algorithm uses
the Gaussian
\begin{equation}
R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}} \, .
\end{equation} 
There are several energy resolution functions available, where by default
{\tt \#standard} is used:
\begin{quote}
{\tt \tb @sigma\_function = \#standard} 
\end{quote}
The energy resolution function {\tt \#standard} is defined by
\begin{equation}
\label{eq:sigma_e}
\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\, ,
\end{equation}
where the parameters $\alpha, \beta$ and $\gamma$ are provided by the user:
\begin{quote}
{\tt \tb @sigma\_e = \{0.15, 0.0, 0.0\}}
\end{quote}
Currently, another possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta},
which only uses the parameter $\alpha$. It is defined by
\begin{equation}
\sigma(E)= \left\{\begin{array}{cl}
 \alpha \cdot \sqrt{1000}^{-1}\,\sqrt{x-8\cdot10^{-4}}\,,&\mathrm{for}\,\, 
x>1.8\cdot10^{-3}\\
5\cdot10^{-5} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
\end{array} \right.
\end{equation}
In the actual implementation of the algorithm,  the sum in \eq~(\ref{eq:algo_one}) is only computed for the $E_j$'s where $K(E_j)$ 
is above a certain threshold, which is by default $10^{-7}$. 
This threshold is defined at the compiling time.

@@

The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
error dimension variable\index{Error dimension}, and the different possibilities are shown in
\Tab~\ref{tab:error_dim}. The error dimensions can be chosen for each rule 
by {\tt @errordim}. For example, 
\begin{quote}
{\tt 
\tb @errordim = 1
}
\end{quote}
In general, it is possible to extend the set of error dimensions or the 
set of possible systematical errors, but this requires C programming, 
compiling and linking with \GLOBES. 
%The interface for user defined error 
%dimensions is described in detail in section~\ref{sec:ud_error_dim}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{table}[t!]
\begin{center}
\begin{tabular}[h]{|c|cccc|c|l|}
\hline
Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
\hline
\hline
0&+&+&+&+&T&Systematics with tilt\\
2&-&-&-&-&-&No systematics\\
4&+&+&+&+&T&Total rates\\
7& $\infty$ &-& $\infty$ &-&-&Spectrum only\\
8&-&-&-&-&-&Total rates, No systematics\\
9&+&+&+&+&C&Systematics with calibration\\
\hline
\end{tabular}
\caption[Table of error dimensions]{\label{tab:error_dim}
Possible values of the error dimensions variable in \GLOBES\ and their meaning. If a parameter is designated with $+$, it will
be marginalized over, and therefore the corresponding error needs to 
have a non-zero value. If the cases with ``total rates'' in the remarks, 
the summation over the bins is performed \emph{before} computing 
the $\chi^2$, \ie, no spectral information is used. 
The error dimension~7 ``spectrum only'' leaves the 
normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
only the spectral information is used. As a consequence,
the settings for the normalization error will be ignored (designated 
with the symbol $\infty$). 
 }
\end{center} 
\end{table} 
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

If you want to have several experiments in one file, separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules and energy resolution
functions. All variables have their scope limited 
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
a consistent treatment of various experiments in one file.

As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include ./file\_1}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include NuFact.gls\\
        \#NEXT\#\\
        include JHFHK.gls
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. 
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments. 
Note that, alternatively, one can 
also load both files successively.

@@

@@

Before we come to the calculation algorithms, it is useful to understand
the general evaluation algorithm. As it is illustrated in \figu{energies}, 
\GLOBES\ uses several levels with respect to the energy ranges:
\begin{description}
\item[Sampling point level]
 This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points. The energy scale is the actual incident neutrino energy $E$. Note that the sampling range should exceed the energy bin range by about the energy resolution error, since the corresponding sampling points have to be evaluated by the algorithm. 
However, this exceeding is only
necessary to avoid aliasing effects, if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
For a manual definition of the sampling points (depends on algorithm), use
\begin{quote}
{\tt
\$sampling\_points = 20\\
\$sampling\_min =          4.0\\
\$sampling\_max =         50.0
}
\end{quote}
for equidistant sampling points. Arbitrarily spaced sampling points can 
be specified with \GLB{\$sampling\_stepsize}
\begin{quote}
{\tt
\$sampling\_stepsize=\{1.0,2.0,3.0,4.0,5.0,...\}
}
\end{quote}

Note that the algorithm expects to 
have at least as many sampling points as energy bins, since otherwise 
the number of (unphysical) sampling constrains the experiment performance.
\item[Bin level] This level is determined by the experiment. Note than energy
bin sizes much smaller than the energy resolution will not improve the results. The energy bin range and the number of energy bins do always have to be specified. For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about the energy calibration error in order to avoid aliasing effects.

@@

As far as the parameterization for the energy resolution function 
$R^c(E,E')$ in \eq~(\ref{eq:kernel}) is concerned, the algorithm uses
the Gaussian
\begin{equation}
R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}} \, .
\end{equation} 
There are several energy resolution functions available, where by default
{\tt \#standard} is used:
\begin{quote}
{\tt \tb @sigma\_function = \#standard} 
\end{quote}
The energy resolution function {\tt \#standard} is defined by
\begin{equation}
\label{eq:sigma_e}
\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\, ,
\end{equation}
where the parameters $\alpha, \beta$ and $\gamma$ are provided by the user:
\begin{quote}
{\tt \tb @sigma\_e = \{0.15, 0.0, 0.0\}}
\end{quote}
Currently, another possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta},
which only uses the parameter $\alpha$. It is defined by
\begin{equation}
\sigma(E)= \left\{\begin{array}{cl}
 \alpha \cdot \sqrt{1000}^{-1}\,\sqrt{x-8\cdot10^{-4}}\,,&\mathrm{for}\,\, 
x>1.8\cdot10^{-3}\\
\alpha\cdot10^{-3} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
\end{array} \right.
\end{equation}
In the actual implementation of the algorithm,  the sum in \eq~(\ref{eq:algo_one}) is only computed for the $E_j$'s where $K(E_j)$ 
is above a certain threshold, which is by default $10^{-7}$. 
This threshold is defined at the compiling time.

@@

The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
error dimension variable\index{Error dimension}, and the different 
possibilities are shown in
\Tab~\ref{tab:error_dim}. Since the error dimension defines the treatement
of systematics it is useful to have define a matched pair of error dimensions
for each rule, where on value describes how the event rate is computed
for \emph{no} systematics and the other one is with systematics 
(see also~\Sec~\ref{sec:systematics}).
The error dimensions for the case of no systematics is set for each rule 
by the value of \GLB{@errordim\_sys\_off}, whereas the error dimension
for systematics on is given by \GLB{@errordim\_sys\_on}. Thus, for example,
the complete error dimension definition could look like  
\begin{quote}
{\tt 
\tb @errordim\_sys\_off = 2\\
\tb @errordim\_sys\_on = 0
}
\end{quote}
It is foreseen to add the possiblity to extend the set 
of error dimensions or the set of possible systematical errors in one
of the next versions of \GLOBES.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{table}[t!]
\begin{center}
\begin{tabular}[h]{|c|cccc|c|l|}
\hline
Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
\hline
\hline
0&+&+&+&+&T&Systematics with tilt\\
2&-&-&-&-&-&No systematics\\
4&+&+&+&+&T&Total rates\\
7& $\infty$ &-& $\infty$ &-&-&Spectrum only\\
8&-&-&-&-&-&Total rates, No systematics\\
9&+&+&+&+&C&Systematics with calibration\\
\hline
\end{tabular}
\caption[Table of error dimensions]{\label{tab:error_dim}
Possible values of the error dimensions variable in \GLOBES\ and their meaning. If a parameter is designated with $+$, it will
be marginalized over, and therefore the corresponding error needs to 
have a non-zero value. If the cases with ``total rates'' in the remarks, 
the summation over the bins is performed \emph{before} computing 
the $\chi^2$, \ie, no spectral information is used. 
The error dimension~7 ``spectrum only'' leaves the 
normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
only the spectral information is used. As a consequence,
the settings for the normalization error will be ignored (designated 
with the symbol $\infty$). 
 }
\end{center} 
\end{table} 
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
